\section{Program}
The program we implemented is a clone of the (at the time of writing) very popular web game 2048\cite{2048}. 

\subsection{Game rules\label{game_rules}}
The game board consists of 4x4 slots.
On game start two random slots will be filled with numbered tiles.
The player must use the directional buttons (SW1 through SW4) to move the tiles.
After each valid move, a new tile will be placed in a random empty tile.
When two tiles of equal value collide they merge into a single tile with value equal to the sum of its' parts.
The game is "won" when the player is able to merge two 1024 tiles into a 2048 tile, although the player may continue the game and try to achieve as high a score as possible.
Game over happens when the board is filled and no move is possible.

\subsubsection{Score}
When two tiles are merged, the resulting value of the merge is added to the current score.

\subsection{Testing the program}
All tests require possession of an EFM32GG-DK3750, a purpose built gamepad connected to the board on GPIO ports A and C and a computer capable of compiling for ARM based platforms and flashing software to the development board.

\subsubsection{Setup}
\begin{enumerate}
    \item   Navigate to the directory \texttt{OSELAS.BSP-EnergyMicro-Gecko}.
    \item   \textt{> make init}\\
	    Make will complain about an error 255, but everything is working as intended.
    \item   \texttt{> make build}\\
	    Again, ignore error 255.
    \item   Shortly after \texttt{make build} has finished the screen on the development board should switch to displaying the game screen.\\
	    The game is now ready to play.
\end{enumerate}

\subsubsection{Playing the game}
\begin{enumerate}
    \item   Use the directional buttons SW1 through SW4 and play the game according to the game rules as described in section \ref{game_rules}.
    \item   When merging tiles, the score displayed on the right of the game board should increase.
    \item   When the current score increases beyond the high score, the high score should increase to equal the current score as well.
    \item   When game over happens the tiles should turn black and display "GAME OVER".
\end{enumerate}

\subsubsection{Restart game}
\begin{enumerate}
    \item   At any point a new game can be started by pressing SW6 twice.
    \item   The game board should reset to contain only two tiles.
    \item   The current score should reset to 0.
    \item   The high score should not reset.
\end{enumerate}
